---
title: "NBA Salary Inequality Over Time"
author: "Jack Meullenet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(reldist)
library(lawstat)
library(DescTools)
library(dplyr)
library(readxl)
```
Getting Gini coefficient for every season:

```{r}
newsals <- read_excel("~/funstuff/nba_ontheside/newsals.xlsx")
newsals$season_start = as.factor(newsals$season_start)

sals2 = newsals %>%
  group_by(season_start) %>%
  summarise(gini = Gini(salary, n = n()))
```

Plotting these over time:

```{r}
gini_over_time2 = ggplot(data = sals2, aes(y = gini, x = season_start, group = 1)) + geom_line() +
  geom_hline(yintercept = .63, linetype = "dashed", col = "red") +
  annotate("text", x = "2012", y = .63, label = "South Africa", vjust = -0.3) +
  geom_hline(yintercept = .48, linetype = "dashed", col = "blue") +
  annotate("text", x = "1992", y = .48, label = "United States", vjust = -0.3) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, vjust = 0.5)) +
  labs(x = "Season Start", y = "Gini Coefficient", title = "NBA Gini Coefficient Over Time")
  
gini_over_time2
```
not too surprising - we have seen a rise in top-heavy teams with veterans taking tiny contracts to "ring chase".could prob verify this going by team, gini by team, how many players in each salary bin etc.

Maybe do some summary stats like std, visualize bell curve. can add these as summary terms for sals2. Could also
be interesting to look at talent concentration on NBA teams, revenues and how revenue sharing affected them.

Overlay with a gini of all stars per team? or just layered histograms/densities

```{r}
allstars <- read_excel("~/funstuff/nba_ontheside/allstars.xlsx")
allstars = allstars %>% select(Year, Player, Team)
allstars$Team = as.factor(allstars$Team)
allstars$Year = as.factor(allstars$Year)

teamrollup = allstars %>%
  group_by(Year, Team) %>%
  summarise(n = n())

ASGini = teamrollup %>%
  group_by(Year) %>%
  summarise(gini = Gini(n, n = n()))

talent_over_time2 = ggplot(data = ASGini, aes(y = gini, x = Year, group = 1)) + geom_line() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, vjust = 0.5)) +
  labs(x = "Season", y = "Gini Coefficient", title = "NBA Gini Coefficients by All-Stars per Team")
talent_over_time2
```

talent gini not great, might need to try summing salaries by team or checking how many minimum players are on title winners?

```{r}
sals3 <- read_excel("~/funstuff/nba_ontheside/sals_w_teams.xlsx")
sals3$player = as.factor(sals3$player)
```